Effects of bariatric surgery and weight loss on resting state functional connectivity

### 2021-02-07 manuscript/draft_rsfmri.Rmd
merged changes + better supplemetary tables

```{r "To do", include=FALSE}
# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
```

@@

```{r figBMIdescr, fig.align='center', out.width="100%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
# eval = TRUE, fig.asp = 0.7, fig.width = 6
readRDS(file = "../report/fig/fig_BMIdescr.rds")
```
<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs. n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs. `r round(tab_age$mean[tab_age$group1=="IG"],2)`). Change in BMI throughout the study is depicted in \@ref(fig:figBMIdescr).
<!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast. They underwent medical assessments including an interview, blood draw and anthropocentric measurements before having a one our break for <!-- standadized??  -->breakfast. MRI scanning was done after further psychological tests (for details see @Prehn_2020).

@@

### Minimal functional preprocessing
The function preprocessing module included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`). In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step. We calculated mean framewise displacement (mFD) from the outputs of the motion correction procedure according to @Power_2012.
For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
<!-- *This part of the analysis was described in the preregistration as "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated."*   -->

@@

#### Functional Quality control
Resting state fMRI data quality control was performed according to the protocol by @Ciric_2018.
As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020]. Thus, we did not exclude any participants based on mFD estimates beyond those who failed the anatomical preprocessing. We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018]. In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mFD.
```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
carpets=c("../report/carpetplotadi006fu2.jpeg","../report/carpetplotadi16fu.jpeg")
knitr::include_graphics(carpets)
```
On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality. Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum). All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
We evaluated mFD-FC relationships by computing a functional connectome based on [@Power_2012]. In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series. Then, we correlated mFD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}. We reported the mFD-FC correlation per preprocessing scheme.
Further, we evaluated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation and report mean, median and standard deviations of mFD, DVARS and number of outliers over all participants.

@@

<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration (*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
We did not perform the analysis as preregistered (i.e. without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mFD (Model 1b).
<!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate. The resulting factorial design <!-- check: flexible factorial? --> contained one regressor for each time point per group.

We did not adjust for baseline BMI as described in the preregistration, but explored the relationship of average and change in BMI in a separate group of models.
As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI. This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC. We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for the logarithm of mFD (Model 2b).
As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and the logarithm of mFD and variability in both measures, along with baseline age and sex (Model 3). Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
<!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen. -->For a better understanding of the unique contribution of average and longitudinal change mFD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject mFD + within-subject mFD$ with nuisance covariates age and sex.

@@

### SwE settings
<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. It therefore accommodates any repeated measurement covariance structures. <!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach. Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images. -->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
In contrast to the parametric estimation, wild bootstrap avoids any parametric distributional assumptions (e.g. random field theory), and unlike permutation testing it is also suitable <!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000). Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$. The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas. <!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom. We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01). Significant clusters are defined as family-wise error - corrected p<0.05. -->

### Aggregated FC analysis
<!-- R analysis of aggregated values -->
We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up. In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and logarithm of mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
We performed model comparison between R1 and R0 models
`R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
`R0 = lmer(agg FC ~ timepoint + group +  age + sex + (1|subj))`
and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
As specified in the preregistration, we repeated the above-mentioned interaction analysis for all three time points. Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and change in mFD.

### Functinal decoding
In an exploratory analysis, we compared the resulting contrast maps with whole-brain activation maps from the [NeuroSynth](https://www.neurosynth.org/) database [@Yarkoni_2011]. We uploaded the contrast images for change BMI and average mFD on [Neurovault](), and applied the decoding classifier. This classifier estimates the similarity of meta-analytic activation maps of +500 search terms with our our contrast maps. We reported the three top terms for both contrasts.

@@

lay <- rbind(c(1,1,2,2),
             c(NA,3,3,NA))
gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)
```

@@

####  Model 1b (with adjustment for age, sex and mFD)
Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.

@@

<!-- Neurovault -->
Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](https://neurovault.org/collections/9426/).

@@

```{r "calc DMN two tp age and sex" ,results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)

@@

```{r "qa check DMN model age sex", eval = FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r "DMN2tp results age sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)

@@

```{r "calc DMN two tp age and sex and mFD", results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)

@@

```{r "qa check DMN model age sex mFD", eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r "qa check Rew model age sex", eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

Reward network connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).

##### BMI and mFD model
To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
<!-- ### DMN -->
<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting mFD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
<!-- gsr -->As before, the effect did not survive when using GSR-corrected data.

<!-- ### Reward -->
<!-- cc -->
For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region (AAL). The peak voxel was classified 45% superior frontal gyrus and 37% supplementary motor area (SMA) in the Harvard-Oxford atlas (see Table \@ref(tab:tableBMIFDmodel)). Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
```
Moreover, FC between the NAcc and one cluster in a more posterior region, i.e. motor cortex, was significantly positively related to average mFD. This result remained similar when using GSR corrected data.
```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
```
###### Functional decoding
Functional decoding of these two activation patterns showed different top association terms. For the [negative BMI change cluster](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula	and inferior frontal were the top terms, while for the [average mFD](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor cortex and
premotor were the meta-analytic activation maps most similar to this contract. The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.

#### Aggregated FC analysis
##### DMN (BMI and mFD model)
```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]

@@

##### Reward network (BMI and mFD model)
```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvwithinFD=anova(R1,R03)
anvmeanFD=anova(R1,R04)

@@

### Effect of mFD (exploratory analysis)
#### Whole-brain analysis
To investigate the unique contributions of the average value of and the longitudinal change in mFD, we constructed two additional models.
First, BMI measures are excluded from the design matrix. Covariance of the mFD and outcome measures is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor. Age and sex were nuisance covariates.

@@

<!-- Neurovault -->
Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average mFD and change mFD will be are published on NeuroVault.

>>>>

```{r "To do", include=FALSE}
# [ ] Upload auf Neurovolt, cf. cobidas reporting guidelines
# [ ] Swap mFD with logmFD ??
```

@@

```{r figBMIdescr, fig.align='center', out.width="50%", fig.cap="Trajectoriy of BMI separately for the bariatric surgery group (BARS) and the waiting-list control group (NBARS); individual trajectories are plotted in transparent, mean trajectories including standard deviations in opaque colors.", echo=FALSE}
# eval = TRUE, fig.asp = 0.7, fig.width = 6
readRDS(file = "../report/fig/fig_BMIdescr.rds")
```
<!-- BASELINE CHARACTERISTICS --> Histograms on baseline characteristics revealed, that patients in the control group did not differ notably from the intervention group regarding BMI and mean FD, and only slightly in distribution of sex and age, the control group had a higher number of male participants (n = `r tab_sex[2,2]` vs. n = `r tab_sex[2,1]`) and a higher mean age (`r round(tab_age$mean[tab_age$group1=="KG"],2)` vs. `r round(tab_age$mean[tab_age$group1=="IG"],2)`). Change in BMI throughout the study is depicted in \@ref(fig:figBMIdescr).
<!-- SURGICAL PROCEDURE --> As intervention, $0$ participants received surgery that combines restrictive and malabsorptive mechanisms (RYGB) and $0$ surgery purely restricting energy intake (VSG: n = $0$, GB: n = $0$).
<!-- STUDY PROCEDURE --> Subjects arrived in the morning (between 07:00 and 12:00) after an overnight fast. They underwent medical assessments including an interview, blood draw and anthropocentric measurements before having a one our break for <!-- standadized??  -->breakfast. MRI scanning was done after further psychological tests (for details see @Prehn_2020).

@@

### Minimal functional preprocessing
The function preprocessing module included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`). In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step.
For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).
<!-- *This part of the analysis was described in the preregistration as "After removing the first four volumes, transforms for motion correction (FSL’s MCFLIRT), fieldmap distortion correction (based on fieldmap and FSL’s FUGUE) and coregistration to the subject’s individual anatomical space (using FREESURFER’s bbregister) are calculated."*   -->

@@

#### Functional Quality control
Resting state fMRI data quality control was performed according to the protocol by @Ciric_2018.
As the expected average head motion was high in this sample, it was difficult to determine an appropriate exclusion threshold which would not lead to the exclusion of a high number of individuals [@Beyer_2020]. Thus, we did not exclude any participants based on mean FD estimates beyond those who failed the anatomical preprocessing. We applied best-practice motion correction (ICA-AROMA and CC, plus optionally GSR) and monitored the reduction of motion-related artifact according to current recommendations [@Ciric_2018]. In addition, for models showing a significant time-by-group interaction, we performed a sensitivity analysis excluding the 10% data sets with highest mean FD.
```{r carpetplot, out.width="50%", out.height="30%", fig.cap="Carpet plot for quality control", fig.subcap = c("Example 1 with strong amount of structured noise", "Example 2 with less amount of structured noise"), fig.show='hold',fig.align='center', echo=FALSE}
carpets=c("../report/carpetplotadi006fu2.jpeg","../report/carpetplotadi16fu.jpeg")
knitr::include_graphics(carpets)
```
On the single-subject level, we used `plot_carpet` from `niworkflows.viz.plots` to generate carpet plots which depict denoising quality. Figure \@ref(fig:carpetplot) shows an exemplary carpet plot with (from top to bottom): percent outliers defined by AFNI, mFD, `OutlierCount`, DVARS (spatial root mean square of the data after temporal differencing),voxelwise timeseries from minimally preprocessed, AROMA non-aggressive and aggressive denoising, CC and CC + GSR data (red: GM voxels, green: WM voxels, blue: ventricle, green: cerebellum). All carpet plots were visually checked and if there was still structured noise in the CC+GSR-denoised functional data (such as black/white stripes, signal dropout), the participant was given a rating of 1 in `QA residuals`.
We evaluated mFD-FC relationships by computing a functional connectome based on [@Power_2012]. In more detail, we used the MNI coordinates of 246 cortical and subcortical spherical seed regions with 5mm radius in `NiftiSpheresMasker`, and calculated the Spearman rank correlation between these time series. Then, we correlated mean FD and QC values for each node across participants, and plotted the resulting mFD-FC correlation values in histogram (see Figure \@ref{exampleFC}. We reported the mFD-FC correlation per preprocessing scheme.
Further, we evaluated the distance-dependence of mFD-FC correlations by correlating the euclidean distance between nodes with the mFD-QC correlation and report mean, median and standard deviations of mFD, DVARS and number of outliers over all participants.

@@

<!-- MODEL SETUP --> In a first group of models, we tested the interaction of time and group $FC = group + time + group*time$ as specified in the preregistration (*We will use the SwE toolbox (implemented in SPM 12.7486 run in MATLAB version > 9.0) to calculate the group-by-timepoint interaction*)
We did not perform the analysis as preregistered (i.e. without covariates and additional sensitivity analyses), but adjusted for baseline age and sex (Model 1a), and baseline age, sex and the logarithm of mean FD (Model 1b).
<!-- DESIGN MATRIX --> In this first model, time was represented as factor because we did not consider the effect of time to be strictly linear, and thus continuous modeling of time, inappropriate. The resulting factorial design <!-- check: flexible factorial? --> contained one regressor for each time point per group.

We did not adjust for baseline BMI as described in the preregistration, but explored the relationship of average and change in BMI in a separate group of models.
As proposed by @Guillaume_2014, we calculated the between- and within subject centered values of BMI. This model, containing average BMI and BMI variability, allowed us to disentangle the differential effects of overall BMI and change in BMI on FC. We first estimated their effects in a model adjusting for baseline age and sex (Model 2a) and then additionally controlling for log mean FD (Model 2b).
As we previously reported correlated change in BMI and mFD in this sample [@Beyer_2019], we explored a refined model including average BMI and log mean FD and variability in both measures, along with baseline age and sex (Model 3). Here, we aimed to see whether any effect of change in BMI would be detectable when adjusting for the change in head motion.
<!-- POST HOC -> können wir nochmal überlegen, ob wir das drinhaben wollen. -->For a better understanding of the unique contribution of average and longitudinal change FD measures, we separately employed a third post-hoc, exploratory model: $FC = between-subject FD + within-subject FD$ with nuisance covariates age and sex.

@@

### SwE settings
<!-- MODEL SPECIFICATION --> The marginal model implemented via the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. It therefore accommodates any repeated measurement covariance structures. <!-- Prereg: During the model specification, we will select the modified sandwich estimator (SwE) approach. Thus, we will define the groups and visits as binary numeric values and subject IDs as numeric values, matching the input order of the images. -->We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size, so heteroscedasticity would be considered.
<!-- INFERENCE --> To ensure robustness of our results, we computed estimates with non-parametric test using wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference.
Type C2 for small sample bias adjustment is applied before the bootstrap resampling as recommended in the SwE manual.
In contrast to the parametric estimation, wild bootstrap avoids any parametric distributional assumptions (e.g. random field theory), and unlike permutation testing it is also suitable <!-- permutations don't work with multilevel data (dependencies)--> for longitudinal data.
Unlike the description in the preregistration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01)$, and $1000$ bootstraps due to required computation time (instead of 5000). Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$. The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005].<!-- Prereg: We will use Type 3 small samples bias adjustment and “Approx” option for estimating the degrees of freedom. We will use the Non-Parametric Wild Bootstrap approach with default parameters, 5000 bootstraps and clusterwise inference (cluster forming threshold of p<0.01). Significant clusters are defined as family-wise error - corrected p<0.05. -->

### Aggregated FC analysis
<!-- R analysis of aggregated values -->
We analyzed the aggregate connectivity values using linear mixed models in the R package lme4 [@Bates_2007].
Following the preregistration, we first investigated the time-by-group interaction between baseline and follow-up. In order to conform with our whole-brain analyses, we adjusted for baseline age, sex (and mFD) in two separate models, instead of performing *sensitivity analysis adjusting for baseline age, sex, average of mean FD of both time points and baseline BMI* (as stated in the preregistration).
We performed model comparison between R1 and R0 models
`R1= lmer(agg FC ~ timepoint*group + age + sex + (1|subj))`
`R0 = lmer(agg FC ~ timepoint + group +  + age + (1|subj))`
and the time-by-group point interaction was considered significant if the model comparison between R1 and Null models showed $p < 0.05$ compared with `anova`.
As specified in the preregistration, we repeated the above-mentioned interaction analysis for all three time points. Again, the interaction was considered significant if a model comparison between R1 and R0 yielded $p < 0.05$.
Further, we performed an exploratory analysis with average BMI and variability in BMI for the aggregated FC, in line with our whole brain analysis, and also included average mFD and variability in mFD.

@@

lay <- rbind(c(1,1,2,2),
             c(1,1,2,2),
             c(NA,3,3,NA),
             c(NA,3,3,NA))
gridExtra::grid.arrange(grobs = figList, layout_matrix = lay)

@@

####  Model 1b (with adjustment for age and sex, meanFD)
Similarly, there were no significant clusters when including mFD. Results did not differ between AROMA+CC and AROMA+CC+GSR preprocessing pipelines.

@@

<!-- Neurovault -->
Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).

@@

```{r "calc DMN two tp, age and sex" ,results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)

@@

```{r "qa check DMN model age,sex", eval = FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r "DMN2tp results age, sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)

@@

```{r "calc DMN two tp, age and sex and mFD", results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)

@@

```{r, "qa check DMN model age,sex, mFD", eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r, "qa check Rew model age,sex", eval=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

NAcc connectivity was not significantly related to neither mean nor change in BMI, for either of the preprocessing schemes.
Results with significant cluster co-activation are depicted in Table \@ref(tab:tableBMImodel).

##### BMI and FD model
To disentangle effects of change in BMI and mFD in this dataset, we split mFD into average mFD and change in mFD. Age and sex were kept as nuisance covariates.
<!-- ### DMN -->
<!-- cc --> Average BMI remained significantly negatively associated with the three clusters above when splitting FD into its components. Results are detailed in Table \@ref(tab:tableBMIFDmodel).
Test statistics only differed slightly from the effect found when using mFD as nuisance covariate.
<!-- gsr -->As before, the effect did not survive when using GSR-corrected data.

<!-- ### Reward -->
<!-- cc -->
For the reward network, we found a significant negative association of change in BMI with FC between NAcc and a cluster in the posterior-medial frontal region, i.e. supplementary motor area (see Table \@ref(tab:tableBMIFDmodel)). Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.028$), also when evaluating the GSR corrected data ($p = 0.038$).
```{r "BMIFDagesexNACCcgnBMIdeact", fig.cap="Significant cluster with a negative Effect of change BMI on NAcc to posterior-medial frontal region, adjusted for age, sex, mean BMI and mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
```
```{r "BMIFDagesexNACCavgFDact", fig.cap="Significant cluster with a positive Effect of average mFD on connectivity between NAcc and motor cortex, adjusted for age, sex, mean BMI, change in BMI and change in mFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
```
```{r "BMIFDagesexPCCccavgBMIdeact2", fig.cap="Negative association of average BMI and FC with PCC/precuneus in AROMA+CC preprocessed data", out.width = "50%"}
knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
```
Moreover, FC between the NAcc and one cluster in a more posterior region, i.e. motor cortex, was significantly positively related to average mFD. This result remained similar when using GSR corrected data.

#### Aggregated FC analysis
```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]

@@

#### Reward network, mean and within variability in BMI and logmFD
```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvwithinFD=anova(R1,R03)
anvmeanFD=anova(R1,R04)

@@

### Effect of FD (exploratory analysis)
#### Whole-brain analysis
To investigate the unique contributions of the average value of and the longitudinal change in mFD, we constructed two additional models.
First, BMI measures are excluded from the design matrix. Covariance of the mFD and outcome measures is investigated by splitting mFD into average mFD and longitudinal mFD change and by inspecting associations with each predictor. Age and sex were nuisance covariates.

@@

<!-- Neurovault -->
Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average FD and change FD will be are published on NeuroVault.

# Discussion
In this preregistered study, we investigated the effects of bariatric surgery on the FC of major resting state brain networks in a longitudinal controlled design. Moreover, we explored the longitudinal relationship of surgery-induced weight loss and FC, and carefully adjusted for head motion by using two efficient denoising schemes and controlling for mean framewise displacement (FD) in group analysis. 
We found no interaction effect of group and timepoint or main effects on whole-brain FC of the PCC and Nacc, which we used to define the resting state networks DMN and reward network, respectively.  Furthermore, we did not find significant interaction or main effects of group and time point on within-network FC. 
In an exploratory model disentangling the effects of average and change in BMI, higher BMI was negatively associated with DMN connectivity for the more lenient denoising scheme. When we additionally adjusted for average and change in head motion, a significant negative relationship between the individuals' change in BMI between the three time points and the connectivity of Nacc with a posterior-medial frontal cluster emerged. This result was significant in both denoising schemes. Functional decoding revealed similarities of the connectivity pattern with frontal, anterior insula and inferior frontal activation patterns. Finally, higher average head motion was associated with increased Nacc connectivity with a cluster in precentral gyrus, close to, yet more posterior than the BMI change cluster.
